Memristor-based recurrent modules for neural computing

### 2023-10-31 paper/circuit.tex
Rearranged figures
Removed color
Fixed tables
added lstm symbol

\section{Design}

@@

\begin{table}[h]
  \centering
  \begin{tabular}{|c|c|}
    \hline
    \rowcolor{gray}
    Real value & Voltage \\
    \hline
    $0$ & $0.9V$ \\
    \hline
    $1$ & $1.0V$ \\
    \hline
    $x$ & $V(x)=\frac{x}{10}+V_{cm}$\\
    \hline
    $real(v)=(v-0.9)\cdot 10$ & $v$\\
    \hline
  \end{tabular}
  \caption{Real/Voltage Conversion Table.}
  \label{tab:valConv}
\end{table}

@@

\begin{figure}[htbp]
  \centering
  \includesvg[width=\columnwidth,pretex=\small]{activation/afCircuit}
  \caption{Activation functions circuit}
  \label{circt:af}
\end{figure}

\begin{table}[htbp]
  \caption{Circuits parameters}
  \label{tab:afPar}
  \centering
  \begin{tabular}{|c|c|c|}
    \hline
    \rowcolor{gray}
    Parameter & Sigmoid & \ac{tanh} \\
    \hline
    $V_1$ & \multicolumn{2}{c|}{$1.1V$}\\
    \hline
    $V_2$ & \multicolumn{2}{c|}{$635mV$}\\
    \hline
    $V_3$ & $0.8V$ & $550mV$\\
    \hline
    $i_{dc}$ & \multicolumn{2}{c|}{$150uA$}\\
    \hline
    $w$ & \multicolumn{2}{c|}{$900nm$}\\
    \hline
    $l$ & \multicolumn{2}{c|}{$60nm$}\\
    \hline
    $R_1$ & \multicolumn{2}{c|}{$5k\Omega$}\\
    \hline
    $R_2$ & \multicolumn{2}{c|}{$10k\Omega$}\\
    \hline
    $R_3$ & $2k\Omega$ & $4k\Omega$\\
    \hline
  \end{tabular}
\end{table}

@@

\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth, pretex=\scriptsize]{activation/afGraph}
  \caption{Input/Output graph of the activation function circuit for both sigmoid and \acs{tanh} functions}
  \label{graph:af}
\end{figure}

@@

\begin{figure}[htbp]
  \centering
  \hspace*{0.8cm}
  \subfloat[Sigmoid symbol]{\includesvg[height=1cm,pretex=\small]{activation/sigmoidSymbol}}%
  \hfill
  \subfloat[\ac{tanh} symbol]{\includesvg[height=1cm,pretex=\small]{activation/tanhSymbol}}%
  \hspace*{0.8cm}
  \caption{Activation functions symbols with the input and output pins on either side depending on the flow of the current for better readability}
  \label{sym:af}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth,pretex=\small]{memcell/memCircuit}
  \caption{Memory cell circuit}
  \label{circt:memcell}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth,pretex=\scriptsize]{memcell/data-loss}
  \caption{Memory conservation in a memory cell with 1 \ac{CMOS} switch vs 2 \ac{CMOS} swicthes}
  \label{fig:memcellLoss}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includesvg[height=1cm,pretex=\small]{memcell/memcellSymbol}
  \caption{Memory cell symbol with the input enable pin (top) and the output enable pin (bottom). The left and right pins are interchangebly the input and ouput for the sake of readability}
  \label{sym:memcell}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth,pretex=\scriptsize]{crossbar/crossbarUse}
  \caption{Circuit of the crossbar array used in the final system ($n_i$, $n_o$, $n_s$)}
  \label{circt:xbar}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth,pretex=\tiny]{crossbar/doubleMem}
  \caption{Simplified circuit of a double memristor per synapse architecture}
  \label{circt:doubleMem}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \begin{tikztimingtable}
    $e_0$ & x2H2L N(A1) L N(A2) 2L N(A3) L N(A4) 2Lx\\
    $e_1$ & x2L2H N(B1) L N(B2) 2L N(B3) L N(B4) 2Lx\\
    $e_i$ & x4L N(C1) L N(C2) 2H N(C3) L N(C4) 2Lx\\
    $e_{n_s-1}$ & x4L N(D1) L N(D2) 2L N(D3) L N(D4) 2Hx\\
    %foo & 2L N(A1)  4H N(A2) L\\
    \extracode
    \node[gap, at={($(A1|-A2)!0.5!(A2)$)}];
    \node[gap, at={($(A3|-A4)!0.5!(A4)$)}];
    \node[gap, at={($(B1|-B2)!0.5!(B2)$)}];
    \node[gap, at={($(B3|-B4)!0.5!(B4)$)}];
    \node[gap, at={($(C1|-C2)!0.5!(C2)$)}];
    \node[gap, at={($(C3|-C4)!0.5!(C4)$)}];
    \node[gap, at={($(D1|-D2)!0.5!(D2)$)}];
    \node[gap, at={($(D3|-D4)!0.5!(D4)$)}];
    \tablerules
    %\draw (0,0) circle (2pt); % Origin
    \begin{pgfonlayer}{background}
      \vertlines[help lines]{0.55,10.55}
      %\vertlines[red]{1.6,5.6,15.6}
      %\vertlines[blue]{3.6,9.6,15.6}
    \end{pgfonlayer}
  \end{tikztimingtable}
  \caption{Enable flags timing for any value of $n_s$ in a single time step}
  \label{tim:serpar}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \includesvg[height=1cm]{crossbar/xbarSymbol}
  \caption{Symbol used for the crossbar array, the input pin is a bus of size $n_i$ and the output pin is a bus of size $n_o$}
  \label{sym:xbar}
\end{figure}

@@

\begin{figure}[h]
  \centering
  \hspace*{1.5cm}
  \subfloat[\ac{opAmp}'s symbol\label{sym:opAmp}]{\includesvg[height=1cm]{models/opAmpSymbol}}%
  \hfill
  \subfloat[Voltage multiplier's symbol\label{sym:voltMult}]{\includesvg[height=1cm]{models/voltMultSymbol}}%
  \hspace*{1.5cm}
  \caption{Symbols used for the verilog-A components}
  \label{sym:models}
\end{figure}

@@

\begin{figure*}[h]
  \centering
  \includesvg[width=\textwidth,pretex=\tiny]{lstm/lstmCircuit}
  \caption{\ac{LSTM} circuit}
  \label{circt:lstm}
\end{figure*}

@@

@@

This section describes the circuit of an encoder \ac{GRU} with an input vector of size $n_i$, a $n_h$ hidden states and $n_{ts}$ time steps. The decoder \ac{GRU} could also be implemented in an analog circuit, but the choice was made to focus on the encoder \ac{GRU}. The \ac{GRU} is by its nature very similar to the \ac{LSTM}. For that reason it has a similar circuit to LSTM circuit (\cref{circt:lstm}). The system is built, once again, using crossbar array (\cref{sec:xbarCircuit}) with ($n_i+n_h+1$,$n_h$, $1$) as parameters.

>>>>

\section{Circuits}

@@

\begin{table}[b]
  \caption{Real/Voltage Conversion Table.}
  \label{tab:valConv}
  \centering
  \begin{tabular}{|c|c|}
    \hline
    \textbf{Real value} & \textbf{Voltage} \\
    \hline
    $0$ & $0.9V$ \\
    \hline
    $1$ & $1.0V$ \\
    \hline
    $x$ & $V(x)=\frac{x}{10}+V_{cm}$\\
    \hline
    $real(v)=(v-0.9)\cdot 10$ & $v$\\
    \hline
  \end{tabular}
\end{table}

@@

\begin{figure}[t]
  \centering
  \includesvg[width=\columnwidth,pretex=\small]{activation/afCircuit}
  \caption{Activation functions circuit}
  \label{circt:af}
\end{figure}

\begin{table}[b]
  \caption{Circuits parameters}
  \label{tab:afPar}
  \centering
  \begin{tabular}{|c|c|c|}
    \hline
    \textbf{Parameter} & \textbf{Sigmoid} & \textbf{\ac{tanh}} \\
    \hline
    $V_1$ & \multicolumn{2}{c|}{$1.1V$}\\
    \hline
    $V_2$ & \multicolumn{2}{c|}{$635mV$}\\
    \hline
    $V_3$ & $0.8V$ & $550mV$\\
    \hline
    $i_{dc}$ & \multicolumn{2}{c|}{$150uA$}\\
    \hline
    $w$ & \multicolumn{2}{c|}{$900nm$}\\
    \hline
    $l$ & \multicolumn{2}{c|}{$60nm$}\\
    \hline
    $R_1$ & \multicolumn{2}{c|}{$5k\Omega$}\\
    \hline
    $R_2$ & \multicolumn{2}{c|}{$10k\Omega$}\\
    \hline
    $R_3$ & $2k\Omega$ & $4k\Omega$\\
    \hline
  \end{tabular}
\end{table}

@@

\begin{figure}[t]
  \centering
  \includesvg[width=\columnwidth, pretex=\scriptsize]{activation/afGraph}
  \caption{Input/Output graph of the activation function circuit for both sigmoid and \acs{tanh} functions}
  \label{graph:af}
\end{figure}

@@

\begin{figure}[b]
  \centering
  \hspace*{0.8cm}
  \subfloat[Sigmoid symbol]{\includesvg[height=1cm,pretex=\small]{activation/sigmoidSymbol}}%
  \hfill
  \subfloat[\ac{tanh} symbol]{\includesvg[height=1cm,pretex=\small]{activation/tanhSymbol}}%
  \hspace*{0.8cm}
  \caption{Activation functions symbols with the input and output pins on either side depending on the flow of the current for better readability}
  \label{sym:af}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includesvg[width=\columnwidth,pretex=\small]{memcell/memCircuit}
  \caption{Memory cell circuit}
  \label{circt:memcell}
\end{figure}

@@

\begin{figure}[b]
  \centering
  \includesvg[width=\columnwidth,pretex=\scriptsize]{memcell/data-loss}
  \caption{Memory conservation in a memory cell with 1 \ac{CMOS} switch vs 2 \ac{CMOS} swicthes}
  \label{fig:memcellLoss}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includesvg[height=1cm,pretex=\small]{memcell/memcellSymbol}
  \caption{Memory cell symbol with the input enable pin (top) and the output enable pin (bottom). The left and right pins are interchangebly the input and ouput for the sake of readability}
  \label{sym:memcell}
\end{figure}

@@

\begin{figure}[b]
  \centering
  \includesvg[width=\columnwidth,pretex=\scriptsize]{crossbar/crossbarUse}
  \caption{Circuit of the crossbar array used in the final system ($n_i$, $n_o$, $n_s$)}
  \label{circt:xbar}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includesvg[width=\columnwidth,pretex=\tiny]{crossbar/doubleMem}
  \caption{Simplified circuit of a double memristor per synapse architecture}
  \label{circt:doubleMem}
\end{figure}

@@

\begin{figure}[b]
  \centering
  \begin{tikztimingtable}
    $e_0$ & x2H2L N(A1) L N(A2) 2L N(A3) L N(A4) 2Lx\\
    $e_1$ & x2L2H N(B1) L N(B2) 2L N(B3) L N(B4) 2Lx\\
    $e_i$ & x4L N(C1) L N(C2) 2H N(C3) L N(C4) 2Lx\\
    $e_{n_s-1}$ & x4L N(D1) L N(D2) 2L N(D3) L N(D4) 2Hx\\
    %foo & 2L N(A1)  4H N(A2) L\\
    \extracode
    \node[gap, at={($(A1|-A2)!0.5!(A2)$)}];
    \node[gap, at={($(A3|-A4)!0.5!(A4)$)}];
    \node[gap, at={($(B1|-B2)!0.5!(B2)$)}];
    \node[gap, at={($(B3|-B4)!0.5!(B4)$)}];
    \node[gap, at={($(C1|-C2)!0.5!(C2)$)}];
    \node[gap, at={($(C3|-C4)!0.5!(C4)$)}];
    \node[gap, at={($(D1|-D2)!0.5!(D2)$)}];
    \node[gap, at={($(D3|-D4)!0.5!(D4)$)}];
    \tablerules
    %\draw (0,0) circle (2pt); % Origin
    \begin{pgfonlayer}{background}
      \vertlines[help lines]{0.55,10.55}
      %\vertlines[red]{1.6,5.6,15.6}
      %\vertlines[blue]{3.6,9.6,15.6}
    \end{pgfonlayer}
  \end{tikztimingtable}
  \caption{Enable flags timing for any value of $n_s$ in a single time step}
  \label{tim:serpar}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \includesvg[height=1cm]{crossbar/xbarSymbol}
  \caption{Symbol used for the crossbar array, the input pin is a bus of size $n_i$ and the output pin is a bus of size $n_o$}
  \label{sym:xbar}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \hspace*{1.5cm}
  \subfloat[\ac{opAmp}'s symbol\label{sym:opAmp}]{\includesvg[height=1cm]{models/opAmpSymbol}}%
  \hfill
  \subfloat[Voltage multiplier's symbol\label{sym:voltMult}]{\includesvg[height=1cm]{models/voltMultSymbol}}%
  \hspace*{1.5cm}
  \caption{Symbols used for the verilog-A components}
  \label{sym:models}
\end{figure}

@@

\begin{figure*}[t]
  \centering
  \includesvg[width=\textwidth,pretex=\tiny]{lstm/lstmCircuit}
  \caption{\ac{LSTM} circuit}
  \label{circt:lstm}
\end{figure*}

@@

The symbol of the \ac{LSTM} circuit can be found in \cref{sym:lstm}. It depends on several parameters, the number of inputs ($n_i$), the number of hidden states ($n_h$), the serial size of the system ($n_s$) and the number of time steps ($n_{ts}$).

\begin{figure}[b]
  \centering
  \includesvg[height=1cm,pretex=\footnotesize]{lstm/lstmSymbol}
  \caption{Symbol used for the \ac{LSTM} circuit}
  \label{sym:lstm}
\end{figure}

@@

This section describes the circuit of an encoder \ac{GRU} with an input vector of size $n_i$, a $n_h$ hidden states and $n_{ts}$ time steps. The decoder \ac{GRU} could also be implemented in an analog circuit, but the choice was made to focus on the encoder \ac{GRU}. The \ac{GRU} is by its nature very similar to the \ac{LSTM}. For that reason it has a similar circuit to LSTM circuit (\cref{circt:lstm}). The system is built, once again, using crossbar array with ($n_i+n_h+1$,$n_h$, $1$) as parameters.
